Choosing a Population Genomics Approach

### 2021-12-18 choosing-population-genetics.Rmd
updating the html choosing pop gen approach

# Population genomic sequencing approaches
Deciding on a sequencing approach for a new project can be challenging. Your choice depends primarily on the questions of interest, but also constraints set by funding, existing genomic resources, sample size, genome characteristics and expertise in the research group. The main categories of population genomic sequencing approaches include (1) High coverage individual whole-genome resequencing (hcWGS) and (2) Low coverage individual WGS (lcWGS), (3).

**Coverage depth** refers to the average number of sequencing reads per region of the genome (distinct from coverage breadth, see [@sims2014sequencing]). With more sequencing reads at a region (higher coverage), errors that arise during sequencing have a smaller impact on accurately identifying variable sites (termed “genotyping”) [@Fuentes-Pardo2017-nu].

### 1) High coverage individual whole-genome resequencing (hcWGS)  {.unlisted .unnumbered}
**Whole-genome resequencing (WGS)** is where whole genomes of multiple individuals are sequenced and mapped to an assembled reference genome in order to compare variable sites between individuals or populations. This approach is considered the is the 'gold standard' for DNA resequencing because it includes all variant types (i.evSNVs, indels, structural variants and CNVs) [@sims2014sequencing].

### 2) Low coverage individual WGS (lcWGS)  {.unlisted .unnumbered}
Low coverage whole genome sequencing is where whole genomes of many individuals are sequenced, but depth of coverage is so low (sequencing depth < 3) that individual genotypes cannot be confidently assigned [@korneliussen2014angsd]. Instead of assigning specific, “hard called" genotypes, lcWGS data requires probabilistic analyses that account for uncertainty about the true genotype and can incorporate uncertainty across all sequenced samples.

### 3) Reduced-representation sequencing {.unlisted .unnumbered}
As the name suggests, this approach aims to sequence a reduced portion of the genome across multiple individuals at moderate to high coverage for variant discovery and genotyping with high confidence. The genome fragments can be selected at random, or through targeted probe assays. Variations of this type of sequencing approach include:

#### RADseq/Genotype-by-sequencing  {.unlisted .unnumbered}
A class of methods involving sequencing a subset of genomic regions randomly distributed throughout the genome, often using restriction enzymes (i.e. restriction site-associated DNA (RADseq)).

#### Targeted capture methods (e.g., whole-exome sequencing (WES))  {.unlisted .unnumbered}
Sequence capture methods use a set of probes designed with a-priori knowledge to focus sequencing effort on a set of hundreds to tens of thousands of specific loci. Several capture approaches for non-model species can utilise other sequencing strategies to design probes corresponding to genomic regions (eg. transcriptome sequencing, ddRAD) (see [@jones2016targeted] for a review).

### 4) Pooled individuals (Poolseq)  {.unlisted .unnumbered}
Sequencing pools of individuals to provide information on genome wide  population  allele  frequencies (see review in [@futschik2010next]).

![Figure 1 from [@Lou2021-me] comparing the distribution of sequencing reads mapped to a reference genome.](images/mec16077-fig-0001-m.jpg)

@@

>>>>

# Population Genomic Sequencing Approaches
Deciding on a sequencing approach for a new project can be challenging. Your choice depends primarily on the questions of interest, but also constraints set by funding, existing genomic resources, sample size, genome characteristics and expertise in the research group.

The main categories of population genomic sequencing approaches include...

## 1) Whole-genome resequencing
Two approaches for whole-genome resequencing are distinguished by their coverage depth. **Coverage depth** refers to the average number of sequencing reads per region of the genome (distinct from coverage breadth, see [@sims2014sequencing]). With more sequencing reads at a region (higher coverage), errors that arise during sequencing have a smaller impact on accurately identifying variable sites (termed “genotyping”) [@Fuentes-Pardo2017-nu].

### High coverage individual whole-genome resequencing (hcWGS)  {.unlisted .unnumbered}
**Whole-genome resequencing (WGS)** is where whole genomes of multiple individuals are sequenced and mapped to an assembled reference genome in order to compare variable sites between individuals or populations.
High-coverage (>20x) sequencing is able to find rare variants with high confidence. This approach is considered the is the 'gold standard' for DNA resequencing because it includes all variant types (i.e SNVs, indels, structural variants and CNVs) [@sims2014sequencing].

### Low coverage individual WGS (lcWGS)  {.unlisted .unnumbered}
**Low coverage whole genome sequencing (lcWGS)** is where whole genomes of many individuals are sequenced, but depth of coverage is so low (<5× per site per individual, on average [@nielsen2011genotype]) that individual genotypes cannot be confidently assigned . Instead of assigning specific, “hard called" genotypes, lcWGS data requires probabilistic analyses that account for uncertainty about the true genotype and can incorporate uncertainty across all sequenced samples [@korneliussen2014angsd]. This approach is useful in cases where the research questions of interest are targeted at the population-level (i.e. allele frequencies, linkage disequilibrium (LD) patterns, etc) [@Lou2021-me].

## 2) Reduced-representation sequencing {.unlisted .unnumbered}
As the name suggests, this approach aims to sequence a reduced portion of the genome across multiple individuals at moderate to high coverage for variant discovery and genotyping with high confidence. The genome fragments can be selected at random, or through targeted probe assays. Variations of this type of sequencing approach include...

### RADseq/Genotype-by-sequencing  {.unlisted .unnumbered}
A class of methods involving sequencing a subset of genomic regions randomly distributed throughout the genome, often using restriction enzymes (i.e. **restriction site-associated DNA (RADseq)**).

### Targeted capture methods  {.unlisted .unnumbered}
Sequence capture methods use a set of probes designed with a-priori knowledge to focus sequencing effort on a set of hundreds to tens of thousands of specific loci. Several capture approaches for non-model species can utilise other sequencing strategies to design probes corresponding to genomic regions (i.e. **Whole-exome sequencing (WES)**, see [@jones2016targeted] for a review, **Expressed Exome Capture Sequencing (eecSeq)**[@puritz2018expressed]).

## 3) Pooled individuals (Poolseq)  {.unlisted .unnumbered}
Sequencing pools of individuals to provide information on genome-wide population allele frequencies (see review in [@futschik2010next]).

![**Figure 1** from [@Lou2021-me] comparing the distribution of sequencing reads mapped to a reference genome.](images/mec16077-fig-0001-m.jpg)

@@

References:
